---
title: "Formative Assessment 4"
author: "Vera Aguila"
date: "2024-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br><br>

The data in Table 5.1 samples of size 50 from a normal distribution, a skewed-right distribution, a skewed-left distribution, and a uniform distribution.

<br><br>

The normal data are female height measurements, the skewed-right data are age at marriage for females, the skewed-left data are obituary data that give the age at death for females, and the uniform data are the amount of cola put into a 12 ounce container by a soft drinks machine.

## Problem 1 
Find the following:
<br>

```{r }
library(moments)

normal <- c(67, 70, 63, 65, 68, 60, 70, 64, 69, 61, 66, 65, 71, 62, 66, 68, 64, 67, 62, 66, 65, 63, 66, 65, 63, 69, 62, 67, 59, 66, 65, 63, 65, 60, 67, 64, 68, 61, 69, 65, 62, 67, 70, 64, 63, 68, 64, 65, 61, 66)

skewed_right <- c(31, 43, 30, 30, 38, 26, 29, 55, 46, 26, 29, 57, 34, 34, 36, 40, 28, 26, 66, 63, 30, 33, 24, 35, 34, 28, 22, 25, 31, 26, 28, 23, 19, 34, 26, 35, 41, 40, 29, 85, 34, 38, 75, 33, 35, 27, 24, 29, 24, 40)

skewed_left <- c(102, 55, 70, 95, 73, 79, 60, 73, 89, 85, 72, 92, 76, 93, 76, 97, 10, 70, 85, 25, 83, 58, 10, 92, 82, 77, 81, 77, 65, 44, 93, 72, 82, 58, 63, 71, 90, 71, 35, 79, 98, 73, 84, 90, 93, 66, 80, 75, 104, 87)

uniform <- c(12.1, 12.1, 12.4, 12.1, 12.1, 12.2, 12.2, 12.2, 11.9, 12.2, 12.3, 12.3, 11.7, 12.3, 12.3, 12.4, 12.4, 12.1, 12.4, 12.4, 12.5, 11.8, 12.5, 12.5, 12.5, 12, 12, 12, 11.9, 12, 11.9, 12.2, 11.9, 11.9, 11.9, 11.8, 11.8, 11.8, 12.5, 11.8, 11.7, 11.7, 11.7, 12.3, 11.7, 11.6, 11.6, 12, 11.6, 11.6)
```

(a) first

```{r, echo=FALSE}
firstN <- sum(normal)/length(normal)
firstSR <- sum(skewed_right)/length(skewed_right)
firstSL <- sum(skewed_left)/length(skewed_left)
firstU <- sum(uniform)/length(uniform)

cat("Normal Distribution First Moments: ")
print(firstN)
cat("Skewed-right Distribution First Moments: ")
print(firstSR)
cat("Skewed-left Distribution First Moments: ")
print(firstSL)
cat("Uniform Distribution First Moments: ")
print(firstU)
``` 


(b) second
```{r, echo=FALSE}
newN <- normal^2
newSR <- skewed_right^2
newSL <- skewed_left^2
newU <- uniform^2

secondN <- sum(newN)/length(newN)
secondSR <- sum(newSR)/length(newSR)
secondSL <- sum(newSL)/length(newSL)
secondU <- sum(newU)/length(newU)

cat("Normal Distribution Second Moments: ")
print(secondN)
cat("Skewed-right Distribution Second Moments: ")
print(secondSR)
cat("Skewed-left Distribution Second Moments: ")
print(secondSL)
cat("Uniform Distribution Second Moments: ")
print(secondU)
``` 
<br>

(c) third
```{r, echo=FALSE}
newN2 <- normal^3
newSR2 <- skewed_right^3
newSL2 <- skewed_left^3
newU2 <- uniform^3

thirdN <- sum(newN2)/length(normal)
thirdSR <- sum(newSR2)/length(skewed_right)
thirdSL <- sum(newSL2)/length(skewed_left)
thirdU <- sum(newU2)/length(uniform)

cat("Normal Distribution Third Moments: ")
print(thirdN)
cat("Skewed-right Distribution Third Moments: ")
print(thirdSR)
cat("Skewed-left Distribution Third Moments: ")
print(thirdSL)
cat("Uniform Distribution Third Moments: ")
print(thirdU)
``` 
<br>

(d) fourth moments for each of the sets of data (normal, skewed-right, skewed-left, uniform)
```{r, echo=FALSE}
newN3 <- normal^4
newSR3 <- skewed_right^4
newSL3 <- skewed_left^4
newU3 <- uniform^4

fourthN <- sum(newN3)/length(normal)
fourthSR <- sum(newSR3)/length(skewed_right)
fourthSL <- sum(newSL3)/length(skewed_left)
fourthU <- sum(newU3)/length(uniform)

cat("Normal Distribution Fourth Moments: ")
print(fourthN)
cat("Skewed-right Distribution Fourth Moments: ")
print(fourthSR)
cat("Skewed-left Distribution Fourth Moments:")
print(fourthSL)
cat("Uniform Distribution Fourth Moments:")
print(fourthU)
``` 

<br><br>

## Problem 2
Find the following:
<br>

(a) first
```{r, echo=FALSE}
newNm <- normal-mean(normal)
newSRm <- skewed_right-mean(skewed_right)
newSLm <- skewed_left-mean(skewed_left)
newUm <- uniform-mean(uniform)

firstNm <- 0
firstSRm <- sum(newSRm)/length(skewed_right)
firstSLm <- sum(newSLm)/length(skewed_left)
firstUm <- sum(newUm)/length(uniform)

cat("Normal Distribution First Moments about the mean: ")
print(firstNm)
cat("Skewed-right Distribution First Moments about the mean: ")
print(firstSRm)
cat("Skewed-left Distribution First Moments about the mean: ")
print(firstSLm)
cat("Uniform Distribution First Moments about the mean: ")
print(firstUm)

``` 
<br>

(b) second
```{r, echo=FALSE}
secondNm <- sum((normal-firstN)^2)/length(normal)
secondSRm <- sum((skewed_right-firstSR)^2)/length(skewed_right)
secondSLm <- sum((skewed_left-firstSL)^2)/length(skewed_left)
secondUm <- sum((uniform-firstU)^2)/length(uniform)

cat("Normal Distribution Second Moments about the mean: ")
print(secondNm)
cat("Skewed-right Distribution Second Moments about the mean: ")
print(secondSRm)
cat("Skewed-left Distribution Second Moments about the mean: ")
print(secondSLm)
cat("Uniform Distribution Second Moments about the mean:\n")
print(secondUm)

``` 
<br>

(c) third

```{r, echo=FALSE}
thirdNm3 <- sum((normal-firstN)^3)/length(normal)
thirdSRm3 <- sum((skewed_right-firstSR)^3)/length(skewed_right)
thirdSLm3 <- sum((skewed_left-firstSL)^3)/length(skewed_left)
thirdUm3 <- sum((uniform-firstU)^3)/length(uniform)

cat("Normal Distribution Third Moments about the mean: ")
print(thirdNm3)
cat("Skewed-right Distribution Third Moments about the mean: ")
print(thirdSRm3)
cat("Skewed-left Distribution Third Moments about the mean: ")
print(thirdSLm3)
cat("Uniform Distribution Third Moments about the mean: ")
print(thirdUm3)
``` 

<br>

(d) fourth moments about the mean for each of the sets of data (normal, skewed-right, skewed-left, uniform)

```{r, echo=FALSE}
fourthNm4 <- sum((normal-firstN)^4)/length(newN)
fourthSRm4 <- sum((skewed_right-firstSR)^4)/length(newSR)
fourthSLm4 <- sum((skewed_left-firstSL)^4)/length(newSL)
fourthUm4 <- sum((uniform-firstU)^4)/length(newU)

cat("Normal Distribution Fourth Moments about the mean: ")
print(fourthNm4)
cat("Skewed-right Distribution Fourth Moments about the mean: ")
print(fourthSRm4)
cat("Skewed-left Distribution Fourth Moments about the mean: ")
print(fourthSLm4)
cat("Uniform Distribution Fourth Moments about the mean: ")
print(fourthUm4)
``` 


<br><br>

## Problem 3
Find the following:
<br>

(a) first

```{r, echo=FALSE}
firstN2 <- sum(normal-75)/length(normal)

cat("Normal Distribution First Moments about the number 75:")
print(firstN2)
``` 
<br>

(b) second

```{r, echo=FALSE}
secondN2 <- sum((normal-75)^2)/length(normal)

cat("Normal Distribution Second Moments about the number 75:")
print(secondN2)
``` 
<br>

(c) third

```{r, echo=FALSE}
thirdN2 <- sum((normal-75)^3)/length(normal)

cat("Normal Distribution Third Moments about the number 75:")
print(thirdN2)
``` 
<br>

(d) fourth moments about the number 75 for the set of female height measurements
```{r, echo=FALSE}
fourthN2 <- sum((normal-75)^4)/length(normal)

cat("Normal Distribution Fourth Moments about the number 75:")
print(fourthN2)
``` 

<br><br>

## Problem 4
Using the results of items 2 and 3 for the set of female height measurements, verify the relations between the moments
<br>
```{r}
m_prime1 = firstN2
m_prime2 = secondN2
m_prime3 = thirdN2
m_prime4 = fourthN2
``` 
<br>

(a) $m_2 = m'_2 - m'^2_1$
```{r}
m2 <- m_prime2 - ((m_prime1)^2)

paste("m2 = ",m2)
``` 

<br>

(b) $m_3 = m'_3 - 3m'_1m'_2 +2m'^3_1$
```{r}
m3 <- m_prime3 - (3*m_prime1*m_prime2) + (2*((m_prime1)^3))

paste("m3 = ",m3)
``` 
<br>

(c) $m_4 = m'_4 - 4m'_1m'_3 + 6m'^2_1m'_2 - 3m'^4_1$
```{r}
m4 <- m_prime4 - (4*m_prime1*m_prime3) + (6*((m_prime1)^2)*m_prime2) - (3*((m_prime1)^4))

paste("m4 = ",m4)
``` 

<br><br>



